### 2024-05-01 Paper/paper.tex

\captionsetup{justification=center,singlelinecheck=false}

@@

    \begin{abstract}
		{\small
        In many industries, including hospitals, nonprofits make up the majority of firms in the market. Understanding the objectives and behaviors of these firms is important when creating incentives. However, nonprofit firms have diverse goals and missions that make understanding their objectives difficult. For hospitals in particular, there is mixed evidence on whether nonprofit hospitals differ meaningfully from for-profit hospitals. In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behaves to for-profits. I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the existence of clinical training on a hospital executive team affects hospital behavior. In particular, I leverage several pay-for-performance initiatives in the US in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives. I find that while readmissions decreased on average for all types of hospitals, the response was larger for for-profits and nonprofits without clinically trained executives. This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put higher weight on societal outcomes than those without clinically trained executives.
		} 
	\end{abstract}

@@

    Second, many researchers have documented a correlation between executive backgrounds, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with choices of female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}). Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}). Due to data limitations, many of these studies focus on large, publicly traded firms in the US. However, one study uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). \textcolor{red}{(add paper about Chile hospital executives)}. I contribute to this literature first by investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.

    Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care. The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}). The author finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}). Importantly, around 40\% of the decrease in readmissions is due to selective patient practices. This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship. This paper addresses a necessary first step in considering whether leadership compositions could affect likelihood of gaming. \textcolor{red}{(add papers about other program too)}

    \section{Institutional Setting}

    \subsection{Hospital Readmissions Reduction Program}\label{sec:hrrp}

@@

    \subsection{Hospital Value-Based Purchasing Program}

    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight. The board is not in charge of day-to-day operations, but the general direction of the firm. They select members of the executive team to carry out day-to-day operations. While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive. An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team. Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc. I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.

@@

    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of executives tied to the hospital from 2009-2015. I gather this data from each hospital's publicly available Tax Form 990, which every sufficiently large nonprofit files each year in the US. To my knowledge, this is the first large scale gathering of leadership names from these forms.

    The historical tax forms for all nonprofits are housed by ProPublica.\footnote{https://projects.propublica.org/nonprofits/} I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are around 3,000 firms left. Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly availble sources. No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets. After limiting to only general, nonprofit hospitals that I can link to the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.

@@

@@

    \begin{equation}
    change_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
    \end{equation}

    \begin{equation}
    change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

>>>>

\captionsetup{justification=centering,singlelinecheck=false}

@@

    \begin{abstract}
		{\small
        In many industries, nonprofits make up the majority of firms in the market. Understanding the objectives and behaviors of these firms is important when creating incentives. However, nonprofit firms have goals and missions that make understanding their objectives difficult. For hospitals in particular, there is mixed evidence on whether nonprofit hospitals differ meaningfully from for-profit hospitals. In this paper, I investigate whether leadership team composition of nonprofit hospitals affects how similarly they behave to for-profits. I use a novel data set on nonprofit hospital executives in the US from 2009-2015 to investigate whether the presence of clinically trained executives affects hospital behavior. In particular, I leverage two pay-for-performance initiatives in the US enacted in 2012, and estimate the difference in response between for-profits, nonprofits, and nonprofits with clinically trained executives. I find that while readmissions decreased on average for all types of hospitals, the response was larger for for-profits and nonprofits without clinically trained executives. This is consistent with theoretical predictions under the assumption that nonprofits with clinically trained executives put greater weight on societal outcomes than those without clinically trained executives.
		} 
	\end{abstract}

@@

    Second, many researchers have documented a correlation between executive backgrounds, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with choices of female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}, \cite{adams2018director}; \cite{frydman2019rising}). Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. However, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals and Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers (\cite{otero2022managers}). I contribute to this literature by looking at the executive team as a whole rather than just CEOs, investigating a context which is not largely studied but is policy relevant, nonprofit US hospitals, and considering a unique executive characteristic pertinent to that context.

    Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care. The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), who finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}). Importantly, around 40\% of the decrease in readmissions is due to selective patient practices. This raises important questions about how to incentivize improvement in quality of care without also incentivizing gamesmanship. Research on HVBP generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}). This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.

    \section{Setting}

    Nonprofits are governed by a board of directors, whose role is to set high level goals and strategies, and provide general oversight. The board is not in charge of day-to-day operations, but the broad direction of the firm. They select executive team members to carry out day-to-day operations. While hospital executives typically do specialize by getting degree(s) in healthcare management or an MBA specific to healthcare, they do not often have clinical training. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive. An executive team usually consists of at least a CEO and CFO, but every nonprofit is different in exactly how they structure their executive team. Some nonprofits hire Chief Medical Officers, Chief Quality Officers, executive directors, vice presidents of various departments, etc. I consider all of these employees as part of the day-to-day management of the hospital and include them as a part of the executive team.

    \subsection{Pay-for-Performance Programs}\label{sec:hrrp}

@@

@@

    To characterize hospitals as having clinically trained executives or not, I construct a novel data set of nonprofit hospitals in the US that contains names and select characteristics of all executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.

    The historical tax forms for all nonprofits are housed by ProPublica.\footnote{https://projects.propublica.org/nonprofits/} I use the NonProfit Explorer API to extract the Employee Identification Numbers (EIN) of all nonprofits classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are around 3,000 firms left. Importantly, I need to be able to link hospitals in the tax forms to other hospital characteristics from publicly available sources. No official crosswalk between EINs and AHA IDs or Medicare numbers exists, so I must rely on matching by location and name of hospital in both data sets. After limiting to only general, nonprofit hospitals that I can link to the AHA survey, I confidently match approximately 1200 EINs to an AHA identification number based on exact name matches within the same state.

@@

    \begin{equation}
    \label{eq:decomp}
    y_{ht} = \beta (no\_MD\_exec_h \times post\_programs_t) + \alpha_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

@@

    \begin{equation}\label{eq:change1}
    change_{ht} = \sum_{j=2011}^{2014}\beta_j\mathbf{1}\{t=j\} + \alpha_h + \epsilon_{ht},
    \end{equation}

    \begin{equation}\label{eq:change2}
    change_{ht} = \beta(program\_exposed_{h} \times post\_2011)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

### 2024-05-01 Tables/forprofit_readmort_synth.tex
Update tables AGAIN

\caption{\label{tab:forprofit_readmort_synth}Comparison to For-Profit Readmission and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{l>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}>{\centering\arraybackslash}p{.2textwidth}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For Profit x Post Programs & -0.07 & -0.32$^{*}$ & -0.03 & -0.01 & -0.17 & 0.02\\
 & (0.04) & (0.12) & (0.06) & (0.04) & (0.11) & (0.06)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
\addlinespace
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
 &  &  &  &  &  & \\
Observations & 15,585 & 2,900 & 4,705 & 15,505 & 2,895 & 4,690\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}

>>>>

\caption{\label{tab:forprofit_readmort_synth}Comparison to For-Profit Readmission and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For-Profit x Post Programs & -0.07 & -0.32$^{**}$ & -0.03 & -0.01 & -0.17 & 0.02\\
 & (0.04) & (0.11) & (0.06) & (0.04) & (0.11) & (0.06)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
 &  &  &  &  &  & \\
Observations & 15,585 & 2,900 & 4,705 & 15,505 & 2,895 & 4,690\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}
